---
title: "glm_validation_and_benchmarking"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{glm_validation_and_benchmarking}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Introduction

This vignette validates the correctness and computational efficiency of the custom my_glm() function, which implements the Iteratively Reweighted Least Squares (IRLS) algorithm for binary logistic regression.
We compare its output and runtime against the base R glm() function.
```{r setup}
library(glmModel)
library(bench)    # For efficiency checking
library(stats)    # Basic R function
set.seed(625)
```

We simulate data with three predictors to make the test more realistic.
```{r}
# Simulated data with an appropriate size to validate  (N=200)
n <- 200
X1 <- rnorm(n)
X2 <- rnorm(n)
X3 <- sample(0:1, n, replace = TRUE)

LP <- 0.5 + 1.2 * X1 - 0.8 * X2 + 0.6 * X3
Prob <- 1 / (1 + exp(-LP))
Y <- rbinom(n, 1, Prob)

df_small <- data.frame(Y, X1, X2, X3)
```

2. Correctness Verification

We first fit the model on the small dataset to ensure the coefficients estimated by my_glm() are numerically equivalent to those from glm() within a tolerance 10^(-4)
```{r correctness_check}
# Fit with custom IRLS

my_model <- my_glm(Y ~ X1 + X2 + X3, data = df_small)

# Fit with base R glm

base_model <- glm(Y ~ X1 + X2 + X3, data = df_small, family = binomial(link = "logit"))

# Print summaries

summary(my_model)
summary(base_model)

# Compare coefficients

is_equal <- all.equal(
as.vector(my_model$coefficients),
as.vector(coef(base_model)),
tolerance = 1e-4
)

cat("\nCoefficient difference can be tolerated:", is_equal, "\n")
stopifnot(isTRUE(is_equal))
```
The output confirms that the coefficient estimates from my_glm are numerically equivalent to those from the base R glm() function, verifying the correct implementation of my IRLS algorithm.


3. Computational Efficiency Benchmarking

We now test computational speed on a large dataset to compare the pure R implementation with the optimized C-level glm().
```{r efficiency_data}
# Simulated data with size 10000
N_large <- 10000
X1_large <- rnorm(N_large)
X2_large <- rnorm(N_large)
X3_large <- sample(0:1, N_large, replace = TRUE)

LP_large <- 0.5 + 1.2 * X1_large - 0.8 * X2_large + 0.6 * X3_large
Prob_large <- 1 / (1 + exp(-LP_large))
Y_large <- rbinom(N_large, 1, Prob_large)

df_large <- data.frame(Y = Y_large, X1 = X1_large, X2 = X2_large, X3 = X3_large)

## Benchmarking
bench_results <- bench::mark(
  # My Rcpp optimized version (my_glm)
  MyGLM_Optimized = my_glm(Y ~ X1 + X2 + X3, data = df_large, max_iter = 20),
  
  # Base R 的 glm
  BaseGLM = glm(Y ~ X1 + X2 + X3, data = df_large, family = binomial(link = "logit")),
  
  iterations = 5, 
  check = FALSE   
)

print(bench_results)
```

4. Conclusion

The coefficient estimates from my_glm() are numerically consistent with glm(). I optimized my code using Rcpp for IRLS part and the runtime for my optimized version code is quicker than R's basic glm function. This confirms that my implementation and optimization of the IRLS algorithm for logistic regression is both correct and functional.
